Crossing Origins by Crossing Formats

### 2012-11-29 main.tex
Updates to Vulnerabilities and attacks

\subsection{Crossing origins}

@@

Other examples of problematic elements include the object- and embed-tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plugins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

\subsection{The object- and embed-tags}
\todo{consider renaming to "Embedded PDF documents", and describe the plug-in behaviour, origin of content, type handling, cookies?}
%The object-tag allows the developer to specify which format 
%the embedded content should be interpreted as.

\todo{consider moving to section "Crossing origins"}
The object- and embed-tags are used to embed content, which requires a browser plug-in
to run, in a web page. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the object tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served.

\todo{consider revising}
As it turns out, various browser plug-ins deal with the situation differently. The Chrome browser
built-in PDF reader refuses to render content that was served with an inappropriate 
MIME-type. The Java plug-in used to ignore the MIME-type, but after several 
vulnerabilities were disclosed, this behavior was changed.

In the case of the Adobe Reader plug-in, it completely disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF. If the 
Chrome browser is configured to use the Adobe Reader plug-in, the behavior will 
be the same as in other browsers and the content will be interpreted as PDF.

\subsection{PDF communication primitives}

% Cross-origin leaks
To leak information across origins, a communication channel is 
required between the attacker and the target. PDF provides the necessary means 
to mediate this information leakage. The flow of information can be divided into 
two steps, extracting information from the target and delivering it to the attacker.
Each requiring its own communication primitives.

When embedded in a web page, many of the PDF communication primitives utilize
the underlying browser to make requests by navigating the browser to the target 
URL. This will unload the current page in which the document is embedded. Such 
primitives can be used in Cross-Site Request Forgery (CSRF) attacks, but would 
defeat the purpose of requesting information as the document cannot process the 
response.

\subsubsection{The target}
There are two primitives for a PDF to communicate with the server it originated from, 
that allow it to read the response; from JavaScript using XML external entities 
or from an embedded flash file.

\paragraph{XML}
The PDF JavaScript API includes a method to parse an XML document, called 
XMLData.parse. The XML document being processed may rely on external entities, 
which the XML parser will request and include in the document. The request is 
bound to the origin of the document. As the response 
is included in the document, it has to be valid, well-formed XML. This puts a 
restriction on the content that can be requested using this method. Considering 
that HTML pages are rarely well-formed this method may be too restricted.
Listing~\ref{lst:xml-pdf} contains an example that uses XML external entities.

@@

\subsubsection{The attacker}
There are a number of ways to communicate with the attacker, either directly 
via the embedding page or through the attackers server. While it is not To prevent the document from unloading

\paragraph{PostMessage}
The PDF plugin includes an API for communicating directly with the page in 
which it is embedded. The document can send messages to the page using 
the hostContainer.postMessage method.

\paragraph{Hash transport}

\paragraph{Embedded Flash}
Cross-origin communication in Flash is restricted by Cross-Origin Resource Sharing. 
As mentioned, the embedded Flash file inherits the origin of the document.
However, this is not a restriction, since an attacker can set a cross-origin policy 
that allow the request.

@@

>>>>

\subsection{Crossing origins}
\label{sec:cross-origin}

@@

Other examples of problematic elements include the object- and embed-tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the object tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plugins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

\subsection{Embedded PDF documents}
\todo{Not sure how well it fits here}
When a PDF document is embedded in a web page, the corresponding plugin 
is executed to render the content. Recall from Section~\ref{sec:cross-origin} 
that the plug-in is selected based 
on the MIME-type, either supplied by the server or in the type attribute, and 
that the browser rely on the plug-in to handle the situation when the MIME-type 
supplied by the server is in appropriate for the plug-in.
In the case of the Adobe Reader plug-in, it disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF.

\subsection{PDF communication methods}

The embedded document will retain the origin from which it was served. Any 
request issued from the document will include any cookies available in the 
browser for the target domain. This provides an attacker with the means to 
perform actions in the name of an authenticated victim, i.e. Cross-Site Request 
Forgery (CSRF). If the attacker can extract information, such as a CSRF-token, 
the protection mechanism is defeated. \todo{this is describing another attack scenario, expand?}

When embedded in a web page, many of the PDF communication methods utilize
the underlying browser to make requests by navigating the browser to the target 
URL. This will unload the current page in which the document is embedded. Page 
navigation is allowed wtih respect to the SOP, since the response is not visible 
to the document. Such methods can be used in CSRF attacks, but would 
defeat the purpose of requesting information as the document cannot process the 
response.

% Cross-origin leaks
To leak information across origins, a communication channel is 
required between the attacker and the target. PDF provides the necessary means 
to mediate this information leakage. The flow of information can be divided into 
two steps, extracting information from the target and delivering it to the attacker.
Each requiring its own communication methods.

\subsubsection{The target}
There are two methods for a PDF to communicate with the server it originated from, 
that allow it to read the response; from JavaScript using XML external entities 
or from an embedded flash file.

\paragraph{XML}
The PDF JavaScript API includes a method to parse an XML document, called 
XMLData.parse. The XML document being processed may rely on external entities, 
which the XML parser will request and include in the document. The request is 
bound to the origin of the document. As the response 
is included in the XML, the result has to be well-formed XML. This puts a 
restriction on the content that can be requested using this method. Considering 
that HTML pages are rarely well-formed this method may be too restricted in practice.
Listing~\ref{lst:xml-pdf} contains an example that uses XML external entities.

@@

\subsubsection{The attacker}
There are a number of ways to communicate with the attacker, either directly 
via the embedding page or through the attackers server. As mentioned previously, 
many methods trigger page navigation in the browser. This will result in a 
noticeable change in the victims browser. While it fulfills the purpose of 
leaking information, an attacker can use more covert methods of communication
to refrain from alerting the victim.

\paragraph{Fragment identifier}
The methods that navigate the browser can still be used for a somewhat covert 
communication with the attacker. Navigating to the same page with a different 
fragment identifier will not reload the page. Embedding the information in the fragment 
identifier of the URL of the embedding website, creates a one-way communication 
channel from the document to the page. Since the fragment identifier is visible 
in the address field of the browser, an alert user would notice it changing.

\paragraph{PostMessage}
The PDF plugin includes an API for communicating directly with the page in 
which it is embedded. The document can send messages to the page using 
the hostContainer.postMessage method. The page registers a handler to receive 
the event and is also able to post messages back to the document.
This two way communication is not visible to the victim.

\paragraph{Embedded Flash}
Cross-origin communication in Flash is restricted by the Same-Origin Policy,
but also support cross-origin communication via Cross-Origin Resource Sharing 
(CORS). Using CORS, a web server can relax the SOP to allow access to specified content.
As mentioned, the embedded Flash file inherits the origin of the document.
However, this is not a restriction, since an attacker can set a cross-origin policy 
that allow the request. This allows for a two way communication that is not 
visible to the victim.

@@

\todo{see if this makes sense here}
As it turns out, various browser plug-ins deal with the situation differently. The Chrome browser
built-in PDF reader refuses to render content that was served with an inappropriate 
MIME-type. If the 
Chrome browser is configured to use the Adobe Reader plug-in, the behavior will 
be the same as in other browsers and the content will be interpreted as PDF.
The Java plug-in used to ignore the MIME-type, but after several 
vulnerabilities were disclosed, this behavior was changed.
